---
title: "Merge IC cores"
author: "K Todd-Brown"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(readxl)
library(tidyverse)

source('../R/read_ReynoldsData.R')
source('../R/read_SiteTextureData.R')
source('../R/read_PattonData.R')
source('../R/read_RyanData.R')

dataDir <- '~/Documents/Professional/Datasets/Renyolds-CZO/data'
```
# Check with Kitty Lohse

1) Do the texture derivations match well?

2) Is there a raster map of the site for the lat/lon plots?

3) Add collector/field notebook/collection date and other metadata locked in KL's head :D.

4) What does BDL in the Patton_DOI_ReynoldsCreek data file mean?

5) Check column names for RyanWill data?

# Read and mash data

## Reynolds Data
```{r setReads, warning=FALSE}
ReynoldReads <- read_ReynoldsData(dataDir=dataDir)

site.df <- ReynoldReads$site
pit.df <- ReynoldReads$pit
mass.df <- ReynoldReads$mass
inorganicC.df <- ReynoldReads$inorganicC

write_excel_csv(site.df, file.path('..', 'temp', 'Reynolds_data', 'site.csv'))
write_excel_csv(pit.df, file.path('..', 'temp', 'Reynolds_data', 'pit.csv'))
write_excel_csv(mass.df, file.path('..', 'temp', 'Reynolds_data', 'mass.csv'))
write_excel_csv(inorganicC.df, file.path('..', 'temp', 'Reynolds_data', 'inorganicC.csv'))
```
## Texture data
```{r readTexture}
texture.df <- read_SiteTextureData(dataDir=dataDir) 
write_excel_csv(texture.df, file.path('..', 'temp', 'Site_texture', 'texture.csv'))
```

```{r readPatton}
patton.ls <- read_PattonData(dataDir=dataDir)
meta_patton.df <- patton.ls$meta
site_patton.df <- patton.ls$site
pit_patton.df <- patton.ls$soilSamples

write_excel_csv(meta_patton.df, file.path('..', 'temp', 'Patton_data', 'meta.csv'))
write_excel_csv(site_patton.df, file.path('..', 'temp', 'Patton_data', 'site.csv'))
write_excel_csv(pit_patton.df, file.path('..', 'temp', 'Patton_data', 'pit.csv'))
```

```{r readRyan}
ryan.ls <- read_RyanData(dataDir=dataDir)
write_excel_csv(ryan.ls$meta, file.path('..', 'temp', 'Ryan_data', 'meta.csv'))
write_excel_csv(ryan.ls$samples, file.path('..', 'temp', 'Ryan_data', 'samples.csv'))


temp_ryan <- ryan.ls$samples %>% 
  unite(col=sample, S, ID) %>%
  unite(col=core, N, E, remove=FALSE) %>%
  mutate(`depth interval` = gsub('cm', '', `depth interval`)) %>%
  separate(`depth interval`, into=c('Layer top (cm)', 'Layer bottom (cm)'), sep='-') %>%
  mutate(sample = paste0('RyanWill-', sample),
         core = paste0('RyanWill-', core),
         `RTK?` = recode(`RTK?`, 'Y' = 'Real Time Kinematic (GPS)', 'N' = 'Other')) %>%
  rename( 'GPS Method' = 'RTK?', 'Layer thickness (m)'=depth,
          'Nitrogen (mass percent)'='N__1', 
          'Organic carbon (mass percent)' = OC,
          'Coarse fraction (precent)' = '%CF',
          'Sand fraction (percent)' = '%sand',
          'Silt fraction (percent)' = '%silt',
          'Clay fraction (percent)' = '%clay', 
          'Bulk density [fine] (kg m-3)' = 'BD_f',
          'Bulk density [modeled] (kg m-3)' = 'BD_m',
          'CaCO3 present (Y/N)' = 'CaCO3?',
          'Coarse root (percent)' = 'c_root %',
          'Fine root (percent)' = 'f_root %') %>%
  select(-`t_root%`)

```

```{r}
##'Site 3' is currently orphaned.... need to go back to field notebook
print(paste('Cores that do not match, [Site 3] is a known orphen:', setdiff(texture.df$core, site.df$core)))
```
 
## Harmonized data
```{r writeNewWS}
allSite.df <- site.df %>%
  select(-bulk_density) %>% #remove the blanket bulk density
  mutate_at(c('E', 'N', 'Elevation_m', 'Slope_degrees', 'Aspect_degrees'), as.numeric) %>%
  bind_rows(temp_ryan %>% select(core, N, E, `GPS Method`)) %>%
  rename('Collection Date'='sample_date', "Easting (m)"='E', "Northing (m)" ='N',
         "Elevation (m)"='Elevation_m', "Aspect (degrees)" ="Aspect_degrees",
         "Slope (degrees)" = "Slope_degrees") %>%
  mutate('Collection Date' = format(`Collection Date`, '%Y-%m-%d')) %>%
  bind_rows(patton.ls$site %>%
              rename("GPS Method"="Collection Method")%>%
              mutate('Collection Date' = format(`Collection Date`, '%Y-%m-%d'),
                     core=`Soil Pit IGSN`)) %>%
  unique()

write_excel_csv(allSite.df, path=file.path('..', 'temp', 'AllSites.csv'))

temp_pit <- ReynoldReads$pit %>%
  mutate(Depth_cm = gsub("^'", "", Depth_cm)) %>%
  mutate(Notes = if_else(grepl('^[[:alpha:]]', Depth_cm) | grepl('cm: ', Depth_cm),
                         paste(Notes, Depth_cm), Notes)) %>%
  #mutate(Depth_old = Depth_cm) %>%
  mutate(Depth_cm = regmatches(Depth_cm, gregexpr('\\d+ ?- ?\\d+', Depth_cm))) %>%
  separate(Depth_cm, into=c('Layer top (cm)', 'Layer bottom (cm)'), sep='-', remove = !FALSE) %>%
  mutate_at(c('Layer top (cm)', 'Layer bottom (cm)'), as.numeric) %>%
  filter(is.finite(`Layer top (cm)`)) %>%
  dplyr::rename(`Gravel-pit (percent)` = Gravel_percent )

temp_mass <-mass.df %>%
  rename('Layer bottom (cm)' = ID_Depth_cm) %>%
  mutate_at(c('Layer bottom (cm)', 'Fine_mass_g', 'Gravel_mass_g'), as.numeric) %>%
  mutate('Fine fraction (fraction)' = Fine_mass_g / (Fine_mass_g+Gravel_mass_g),
         'Gravel fraction (fraction)' = Gravel_mass_g/ (Fine_mass_g+Gravel_mass_g)) %>%
  select(core, ID_sample, `Layer bottom (cm)`, `Fine fraction (fraction)`, `Gravel fraction (fraction)`)

temp_texture <- texture.df %>%
  mutate('Sand (fraction)' = `Sand (g)`/`Total (by oven) (g)`,
         'Silt (fraction)' = `Silt (g)`/`Total (by oven) (g)`,
         'Clay (fraction)' = `Clay (g)`/`Total (by oven) (g)`) %>%
  mutate(`Depth (cm)` = gsub("^'", "", `Depth (cm)`)) %>%
  separate(`Depth (cm)`, into=c('Layer top (cm)', 'Layer bottom (cm)'), sep='-') %>%
  select(core, ID_sample, starts_with('Layer'), ends_with('(fraction)')) %>%
  mutate_at(vars(starts_with('Layer'), ends_with('(fraction)')), as.numeric)

temp_ic <- inorganicC.df %>%
  mutate(IC_percent = as.numeric(IC_percent),
         'Layer bottom (cm)' = as.numeric(ID_Depth_cm)) %>%
  mutate(ID_fraction = if_else(grepl('G', ID_fraction), 
                               'Inorganic carbon: gravel (percent)', 'Inorganic carbon: fine (percent)')) %>%
  spread(key=ID_fraction, value=IC_percent) %>%
  select(core, ID_sample, `Layer bottom (cm)`, ends_with('(percent)'))

#sample.df <- full_join(temp_pit, temp_texture, by= c("core", "Layer top (cm)", "Layer bottom (cm)")) %>%
#  full_join(temp_mass, by=c('core', "ID_sample", 'Layer bottom (cm)')) %>%
#  full_join(temp_ic, by=c('core', "ID_sample", 'Layer bottom (cm)')) %>%
  #add in temp_ryan and patton.ls
```

# Cluster sites

```{r cluster}

latLon.df <- allSite.df %>% select(`Easting (m)`, `Northing (m)`) %>% unique %>%
  filter(is.finite(`Easting (m)` + `Northing (m)`)) %>% 
  rename(Easting = `Easting (m)`, Northing = `Northing (m)`)

knownSites <-  allSite.df %>% select(site_name, `Easting (m)`, `Northing (m)`) %>%
  filter(is.finite(`Easting (m)` + `Northing (m)`), !is.na(site_name)) %>% unique %>% 
  rename(site_Easting = `Easting (m)`, site_Northing = `Northing (m)`)

temp <- plyr::ddply(knownSites, c('site_name', 'site_Easting', 'site_Northing'), function(xx){
  ans <- latLon.df %>%
    mutate(distance = sqrt((`Easting` - xx$site_Easting)^2 + (`Northing`-xx$site_Northing)^2)) %>%
    filter(distance < 30)
  return(ans)
})

clusterLatLon <- latLon.df %>% full_join(temp, by = c("Easting", "Northing"))

ggplot(clusterLatLon) +
  geom_point(aes(x=Easting, y=Northing, color=is.na(site_name)))
```

# QA/QC
 
## Site.df
```{r}

##List the layers ####
rgdal::ogrListLayers('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/access_roads.kml')
rgdal::ogrListLayers('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/hydrology.kml')
rgdal::ogrListLayers('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/boundary.kml')
#rgdal::ogrListLayers('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/NED_10m/ned_10m/')

##Read in layers ####
roads.sp <- rgdal::readOGR('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/access_roads.kml',
                           layer='RCEW.DBO.RCEW_DBO_RC_Roads')
hydro.sp <- rgdal::readOGR('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/hydrology.kml',
                           layer='Hydrology')
boundry.sp <- rgdal::readOGR('~/Documents/Professional/Datasets/Renyolds-CZO/GIS_dataproducts/boundary.kml',
                             layer='RCEW_Boundary')

ggplot(data=fortify(boundry.sp)) +
  geom_polygon(aes(x=long, y=lat, group=group), color='grey', fill=NA) +
  geom_path(data=fortify(hydro.sp), aes(x=long, y=lat, group=group, order=order), color='blue') +
  geom_path(data=fortify(roads.sp), aes(x=long, y=lat, group=group, order=order), color='black') 
```
```{r}
site.df <- site.df %>% 
  mutate_at(vars(c(E, N, Elevation_m, Slope_degrees, Aspect_degrees)), as.numeric)

ggplot(site.df) +
  #geom_point(aes(x=E, y=N), color='yellow') +
  geom_point(aes(x=E, y=N, color=Elevation_m)) +
  theme_bw()

``` 
 
```{r basicVis}
ggplot(site.df) +
  #geom_point(aes(x=E, y=N), color='yellow') +
  geom_point(aes(x=E, y=N, color=core %in% mass.df$core)) +
  theme_bw()
ggplot(site.df) +
  #geom_point(aes(x=E, y=N), color='yellow') +
  geom_point(aes(x=E, y=N, color=core %in% inorganicC.df$core)) +
  theme_bw()


soil.df <- inorganicC.df %>%
  mutate(ID_fraction = if_else(ID_fraction %in% c('G', 'GRAVEL'),
                               'Gravel_IC_percent', 'Fine_IC_percent')) %>%
  spread(key=ID_fraction, value=IC_percent) %>%
  full_join(mass.df) %>%
  mutate_at(vars(-core, -ID_sample), as.numeric) %>%
  mutate_at(c('core', 'ID_sample'), as.factor) %>%
  mutate(Gravel_IC_percent = case_when(is.na(Gravel_IC_percent) ~ 1e-4, 
                                       Gravel_IC_percent < 1e-4 ~ 1e-4,
                                       TRUE ~ Gravel_IC_percent),
         Fine_IC_percent = case_when(is.na(Fine_IC_percent) ~ 1e-4, 
                                       Fine_IC_percent < 1e-4 ~ 1e-4,
                                       TRUE ~ Fine_IC_percent))
  
  
write_excel_csv(soil.df, file.path('..', 'temp', 'Reynolds_data', 'filled_massIC.csv'))
```
```{r}
plot.df <- soil.df %>%
  mutate(plotGroup = paste(core, ID_sample),
         total_IC_percent = (Fine_IC_percent*Fine_mass_g + Gravel_IC_percent*Gravel_mass_g) / 
                  (Fine_mass_g+Gravel_mass_g)) %>%
  gather(key='fraction', value='IC (percent)', contains('IC'))

ggplot(plot.df) +
  geom_line(aes(x=-ID_Depth_cm, y=`IC (percent)`, group=plotGroup)) +
  coord_flip() +
  facet_wrap(~fraction)

ggplot(mass.df %>%
         mutate_at(vars(c(ID_Depth_cm, contains('mass'))), as.numeric) %>%
         mutate(plotGroup = paste(core, ID_sample),
                gravel_fraction = Gravel_mass_g/(Fine_mass_g + Gravel_mass_g))) +
  geom_line(aes(x=-ID_Depth_cm, y=gravel_fraction, group=plotGroup)) +
  coord_flip()
```
